Reference Tracking Strategies of Deaf Adult Signers in Turkish Sign Language

### 2023-05-02 revised_manuscript.Rmd
model results added
Added the model results to the appendix.

---
editor_options:
  chunk_output_type: console
title: Reference Tracking Strategies of Deaf Adult Signers in Turkish Sign Language
abstract: |
  We investigated the reference tracking strategies among deaf adults with signing deaf parents (DoD) and adult deaf signers with non-signing hearing parents or caregivers (DoH) who had late exposure to Turkish Sign Language (TİD). Consistent with the theories of saliency and referential accessibility, regardless of their acquisiton groups or parental hearing status, signers mainly used nominals and extension classifiers for introducing referents. To maintain a referent from the previous clause, the signers used zero anaphora (e.g., constructed action, agreement and plain verbs). For topic shifts or re-introduced contexts, nominals and pronouns were chiefly favored although we observed very little pronominal use. As for the effect of native acquisition from deaf caregivers, we only report limited over-redundancy for DoH signers who used zero anaphora less compared to DoD especially for introduced and maintained contexts. We conclude that DoH signers are still able to achieve native-like competency in terms of reference tracking in simple narratives but DoD signers utilize the spatial affordances of the visual-spatial modality better than DoH signers only to a certain extent.
keywords: "Referential accessibility, signed narratives, pragmatic competence, delayed language acquisition, Turkish Sign Language"
journal: "Journal of Pragmatics"
date: "`r Sys.Date()`"
bibliography: ./bibliography/references.bib
times: true
numbersections: trueand w
classoption: "review"
csl: ./bibliography/elsevier-harvard.csl
output: rticles::elsevier_article
header-includes:
   - \usepackage{floatrow}
   - \usepackage{hyperref}
   - \usepackage{enumitem}  
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \floatplacement{table}{H}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{times}
   - \usepackage{smartdiagram}
   - \usepackage{ragged2e}
   - \usepackage{tikz}
   - \usepackage{xcolor}
   - \usepackage[utf8]{inputenc}
   - \usepackage[T1]{fontenc}
   - \usepackage{graphicx}
   - \usepackage{./sty/langsci-gb4e}
   - \usepackage{caption}
   - \usepackage{multicol}
   - \setlength{\columnsep}{-2cm}
   - \usepackage{setspace}
---

@@

[^2]: We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2. However, our model did not converge properly and the credible intervals were very wide spread, yielding the results uninformative. We provide the model results as a plot in \ref{section:d}.

@@

@@

df_table_extra$Coefficients %<>% dplyr::recode(`muNominal_Intercept` = "Nom_Intercept",
                                        `muNull_Intercept` = "Null_Intercept",
                                        `muNominal_Group2M1` = "Nom_Late1vsNative",
                                        `muNominal_Group3M2` = "Nom_Late2vsLate1",
                                        `muNominal_Discourse1` = "Nom_Intro",
                                        `muNominal_Discourse2` = "Nom_Maintain",
                                        `muNominal_Group2M1:Discourse1` = 'Nom_L1vsN*Intro',
                                        `muNominal_Group3M2:Discourse1` = 'Nom_L2vsL1*Intro',
                                        `muNominal_Group2M1:Discourse2` = 'Nom_L1vsN*Maintain',
                                        `muNominal_Group3M2:Discourse2` = 'Nom_L2vsL1*Maintain',
                                        `muNull_Group2M1` = "Null_Late1vsNative",
                                        `muNull_Group3M2` = "Null_Late2vsLate1",
                                        `muNull_Discourse1` = "Null_Intro",
                                        `muNull_Discourse2` = "Null_Maintain",
                                        `muNull_Group2M1:Discourse1` = 'Null_L1vsN*Intro',
                                        `muNull_Group3M2:Discourse1` = 'Null_L2vsL1*Intro',
                                        `muNull_Group2M1:Discourse2` = 'Null_L1vsN*Maintain',
                                        `muNull_Group3M2:Discourse2` = 'Null_L2vsL1*Maintain')
df_table_extra %<>% mutate( across(.cols = "Coefficients",~str_replace( ., "_","\\\\_")))

>>>>

---
editor_options:
  chunk_output_type: console
title: Reference Tracking Strategies of Deaf Adult Signers in Turkish Sign Language
abstract: |
  We investigated the reference tracking strategies among deaf adults with signing deaf parents (DoD) and adult deaf signers with non-signing hearing parents or caregivers (DoH) who had late exposure to Turkish Sign Language (TİD). Consistent with the theories of saliency and referential accessibility, regardless of their acquisiton groups or parental hearing status, signers mainly used nominals and extension classifiers for introducing referents. To maintain a referent from the previous clause, the signers used zero anaphora (e.g., constructed action, agreement and plain verbs). For topic shifts or re-introduced contexts, nominals and pronouns were chiefly favored although we observed very little pronominal use. As for the effect of native acquisition from deaf caregivers, we only report limited over-redundancy for DoH signers who used zero anaphora less compared to DoD especially for introduced and maintained contexts. We conclude that DoH signers are still able to achieve native-like competency in terms of reference tracking in simple narratives but DoD signers utilize the spatial affordances of the visual-spatial modality better than DoH signers only to a certain extent.
keywords: "Referential accessibility, signed narratives, pragmatic competence, delayed language acquisition, Turkish Sign Language"
journal: "Journal of Pragmatics"
date: "`r Sys.Date()`"
bibliography: ./bibliography/references.bib
times: true
numbersections: trueand w
classoption: "review"
csl: ./bibliography/elsevier-harvard.csl
output: rticles::elsevier_article
header-includes:
   - \usepackage{floatrow}
   - \usepackage{hyperref}
   - \usepackage{enumitem}  
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \floatplacement{table}{H}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{times}
   - \usepackage{smartdiagram}
   - \usepackage{ragged2e}
   - \usepackage{tikz}
   - \usepackage{xcolor}
   - \usepackage[utf8]{inputenc}
   - \usepackage[T1]{fontenc}
   - \usepackage{graphicx}
   - \usepackage{langsci-gb4e}
   - \usepackage{caption}
   - \usepackage{multicol}
   - \setlength{\columnsep}{-2cm}
   - \usepackage{setspace}
---

@@

[^2]: The regression model results are given in \ref{section:modelResult}. We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2. However, our model did not converge properly and the credible intervals were very wide spread, yielding the results uninformative. We provide the model results as a plot in \ref{section:d}.

@@

# Regression model results for RE Type

\label{section:modelResult}

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
colnames(df_model_1)[1:4] <- c("Coefficients","outerWidth","innerWidth", "pointEstimate")
df_model_1 %<>% mutate(across(.cols = "Coefficients", ~str_replace( ., "_","\\\\_")))
apa_table(df_model_1,
          format = "latex",
          booktabs = T,
          escape = FALSE,
          landscape = FALSE,
          font_size = "small",
          placement = "hbt!",
          caption ="Regression model results for the RE Type (Nominal, Zero, Classifier) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH)"
)
```

@@

df_table_extra$Coefficients %<>% dplyr::recode(`muNominal_Intercept` = "Nom_Intercept",
                                        `muNull_Intercept` = "Null_Intercept",
                                        `muNominal_Group2M1` = "Nom_Late1vsNative",
                                        `muNominal_Group3M2` = "Nom_Late2vsLate1",
                                        `muNominal_Discourse1` = "Nom_Intro",
                                        `muNominal_Discourse2` = "Nom_Maintain",
                                        `muNominal_Group2M1:Discourse1` = 'Nom_L1vsN*Intro',
                                        `muNominal_Group3M2:Discourse1` = 'Nom_L2vsL1*Intro',
                                        `muNominal_Group2M1:Discourse2` = 'Nom_L1vsN*Maintain',
                                        `muNominal_Group3M2:Discourse2` = 'Nom_L2vsL1*Maintain',
                                        `muNull_Group2M1` = "Null_Late1vsNative",
                                        `muNull_Group3M2` = "Null_Late2vsLate1",
                                        `muNull_Discourse1` = "Null_Intro",
                                        `muNull_Discourse2` = "Null_Maintain",
                                        `muNull_Group2M1:Discourse1` = 'Null_L1vsN*Intro',
                                        `muNull_Group3M2:Discourse1` = 'Null_L2vsL1*Intro',
                                        `muNull_Group2M1:Discourse2` = 'Null_L1vsN*Maintain',
                                        `muNull_Group3M2:Discourse2` = 'Null_L2vsL1*Maintain')
df_table_extra %<>% mutate(across(.cols = "Coefficients",~str_replace( ., "_","\\\\_")))
